---
title: 'Is there a need for a new journal metric? Correlations between JCR Impact Factor
  metrics and the Journal Citation Indicator - JCI'
author: 'Daniel Torres-Salinas, Pilar Valderrama-Baca, Wenceslao Arroyo-Machado'
date: '10 April, 2022'
output:
  html_document:
    toc: yes
---

This is the [R Markdown](http://rmarkdown.rstudio.com) Notebook created for the JCI analysis. Notebook and data are available at [GitHub](https://github.com/Wences91/jci_correlations).
```{r packages, warning=FALSE, message=FALSE}
source('functions.R')
library(reldist)
library(gglorenz)
library(ggplot2)
library(GGally)
library(dplyr)
library(ggnewscale)

set.seed(13)
```

First line only can be executed if you have downloaded data from JCR in multiple files, otherwise, use the second line which includes the anonymized JCR data.
```{r read_jcr}
#df <- read_jcr('data/files/')
df <- read.delim('data/jcr.tsv', stringsAsFactors=FALSE, check.names=FALSE)
```

Some descriptive statistics of the dataset.
```{r stats, h}
summary(df$`2020 JCI`)
```

# 1. Distribution analysis
Gini Index is calculated for each JCR indicator.
```{r gini}
gini(df$`2020 JIF`[which(!is.na(df$`2020 JIF`))])
gini(df$`2020 JCI`[which(!is.na(df$`2020 JCI`))])
gini(df$`5 Year JIF` [which(!is.na(df$`5 Year JIF`))])
gini(df$`Immediacy Index`[which(!is.na(df$`Immediacy Index`))])
gini(df$`Eigenfactor`[which(!is.na(df$`Eigenfactor`))])
gini(df$`Total Citations`[which(!is.na(df$`Total Citations`))])
gini(df$`Article Influence Score`[which(!is.na(df$`Article Influence Score`))])
```

A new `data.frame` is created to obtain the Lorenz curve in an easy way.
```{r lorenz_df}
df_den <- data.frame()

for(x in c('2020 JCI','2020 JIF', '5 Year JIF', 'Immediacy Index', 'Eigenfactor', 'Total Citations', 'Article Influence Score')){
  aux <-df[,c('Cat','DB',x)]
  names(aux)[3] <- 'Value'
  aux$Indicator <- x
  df_den <- rbind.data.frame(df_den, aux, stringsAsFactors = FALSE)
}

df_den[which(df_den$Indicator=='2020 JCI'),'Indicator'] <- 'Journal Citation Indicator'
df_den[which(df_den$Indicator=='2020 JIF'),'Indicator'] <- 'Journal Impact Factor'
df_den[which(df_den$Indicator=='5 Year JIF'),'Indicator']  <- '5-Year Journal Impact Factor'
```

The Lorenz curve is obtained.
```{r lorenz_curve, dpi=100, fig.retina=2, fig.height=5, fig.width=9, warning=FALSE}
ggplot(df_den, aes(x=Value, color=Indicator)) +
  stat_lorenz(desc=FALSE, alpha=0.7, size=1) +
  geom_abline(size=0.8) +
  scale_color_manual(values=c('#fb0007','#129176', '#e5c421', '#f46f08', '#4cacd0', '#291b37', '#de8bb7'))+
  theme_light() +
  theme(panel.grid=element_blank(),
        legend.title=element_blank(),
        text=element_text(family='Arial', size=12.5, color='black'),
        axis.text=element_text(color='black', size=12),
        axis.ticks=element_line(color='black'),
        legend.position='bottom',
        panel.border=element_rect(colour='black'),
        strip.background=element_rect(colour='black', fill='black'),
        strip.text=element_text(size=14),
        legend.text=element_text(size=10))+
  guides(fill=guide_colorbar(label.position='bottom',
                               title.position='left', title.vjust=0.85)) +
  labs(y='Cummulative percentage of indicator', x='Cummulative percentage of journals')
```

# 2. Correlation analysis
A new `data.frame` is created for obtaining a matrix of plots.
```{r matrix_plot_df}
m_plot <- df[,c('2020 JCI', '2020 JIF', '5 Year JIF', 'Immediacy Index', 'Eigenfactor', 'Total Citations', 'Article Influence Score', 'DB')]
names(m_plot)[c(1,2,7)] <- c('JCI', 'JIF', 'AIS')
```

The matrix of plots with the correlations and distribution of data is generated. All of them are significant.
```{r matrix_plot, dpi=200, fig.retina=2,  fig.height=9, fig.width=14, warning=FALSE, message=FALSE}
ggpairs(m_plot, columns=1:7,
        upper=list(continuous=wrap('cor', size=5, method='pearson')),
        lower=list(continuous=wrap('points', alpha=0.6)))+
  theme(text=element_text(family='Arial', size=12.5, color='black'),
        axis.text=element_text(color='black', size=11),
        axis.ticks=element_line(color='black'),
        strip.background=element_rect(colour='black', fill='black'),
        strip.text=element_text(size=12, color='white'),
        legend.position='none'
        ) 
```

Pearson correlations between the seven indicators are calculated.
```{r correlation_plots, fig.width=3, warning=FALSE}
df_cor <- data.frame()
for(x in c('2020 JIF', '5 Year JIF', 'Immediacy Index', 'Eigenfactor', 'Total Citations', 'Article Influence Score')){
  aux <-biblio_cor(df, x, '2020 JCI')
  aux$Indicator <- x
  df_cor <- rbind.data.frame(df_cor, aux, stringsAsFactors = FALSE)
}

df_cor <- df_cor[which(!is.na(df_cor$Correlation)),]
```

ESI field are assigned to each Web of Science Category.
```{r esi}
esi <- read.csv2('data/final_map.csv', stringsAsFactors=FALSE)
esi[which(esi$WC=="Women's Studies"), 'WC'] <- 'Womens Studies'
esi$WC <- toupper(esi$WC)

df_cor <- inner_join(df_cor, esi[,c('WC', 'ESI')], by=c('Cat'='WC'))
```

## 2.1. JIF and 5-Year JIF
Firstly, correlations between JCI and JIF and 5-Year JIF are calculated.
```{r correlations, warning=FALSE}
jci_jif <- df_cor[which((df_cor$Indicator %in% c('2020 JIF', '5 Year JIF')) & (df_cor$DB %in% c('SSCI', 'SCIE'))),]

jci_jif[which(jci_jif$DB=='SSCI'),'DB'] <- 'Social Sciences'
jci_jif[which(jci_jif$DB=='SCIE'),'DB'] <- 'Science'

jci_jif[which(jci_jif$Indicator=='2020 JIF'),'Indicator'] <- 'Journal Impact Factor'
jci_jif[which(jci_jif$Indicator=='5 Year JIF'),'Indicator'] <- '5-Year Journal Impact Factor'

jci_jif$Indicator <- factor(jci_jif$Indicator, levels=c('Journal Impact Factor', '5-Year Journal Impact Factor'), ordered=TRUE)
```

There are 13 correlations with a `p-value` less than 0.01 (6 in JIF and 7 in 5-Year JIF).
```{r cor_sign}
dim(jci_jif[which(jci_jif$p > 0.001),])
```


```{r}
mean(jci_jif[which(jci_jif$p < 0.001 & jci_jif$Indicator=='Journal Impact Factor'),'Correlation'])
mean(jci_jif[which(jci_jif$p < 0.001 & jci_jif$Indicator=='Journal Impact Factor' & jci_jif$DB == 'Science'),'Correlation'])
mean(jci_jif[which(jci_jif$p < 0.001 & jci_jif$Indicator=='Journal Impact Factor' & jci_jif$DB == 'Social Sciences'),'Correlation'])
```

Boxplot with correlations greater than 0.6 is generated.
```{r jci_jif_05, dpi=200, fig.retina=2, fig.height=5, fig.width=9, warning=FALSE, message=FALSE}
ggplot(data=jci_jif[which(jci_jif$p < 0.001),], aes(y=Correlation, x=DB, group=DB))+
  geom_boxplot(outlier.shape=NA, width=0.8) +
  geom_jitter(data=jci_jif[which(jci_jif$Indicator=='Journal Impact Factor' & jci_jif$p < 0.001),], aes(fill=Correlation, size=Docs), shape=21, color='black', stroke=0.5, alpha=0.6)+
  scale_fill_gradient(low='white', high='#eb2701', na.value=NA, breaks=c(0.6, 0.8, 1), limits=c(0.6,1))+
  new_scale_fill()+
  geom_jitter(data=jci_jif[which(jci_jif$Indicator=='5-Year Journal Impact Factor' & jci_jif$p < 0.001),], aes(fill=Correlation, size=Docs), shape=21, color='black', stroke=0.5, alpha=0.6)+
  scale_fill_gradient(low='white', high='blue', na.value=NA, breaks=c(0.6, 0.8, 1), limits=c(0.6,1))+
  ylim(c(0.6,1))+
  scale_size_continuous(range=c(1,5), breaks=c(0, 25000,50000, 100000, 125000, 150000, 200000))+
  theme_light()+
  theme(panel.grid=element_blank(),
        text=element_text(family='Arial', size=12.5, color='black'),
        axis.text=element_text(color='black', size=11),
        axis.ticks=element_line(color='black'),
        legend.position='none',
        legend.box='vertical',
        panel.border=element_rect(colour='black'),
        strip.background=element_rect(colour='black', fill='black'),
        strip.text = element_text(size=14))+
  guides(fill=guide_colorbar(label.position='bottom',
                             title.position='left', title.vjust=0.85)) +
  labs(x='', y='Correlation with JCI')+
  facet_wrap(~Indicator)
```

The complete boxplot does not vary much, only 7 more categories are included.
```{r jci_jif, dpi = 200, fig.retina=2, fig.height=5, fig.width=9, warning=FALSE, message=FALSE}
ggplot(data=jci_jif, aes(y=Correlation, x=DB, group=DB))+
  geom_boxplot(outlier.shape=NA, width=0.8) +
  geom_jitter(data=jci_jif[which(jci_jif$Indicator=='Journal Impact Factor'),], aes(fill=Correlation, size=Docs), shape=21, color='black', stroke=0.5, alpha=0.6)+
  scale_fill_gradient(low='white', high='#eb2701', na.value=NA, breaks=c(0.5, 0.75, 1), limits=c(0.5,1))+
  new_scale_fill()+
  geom_jitter(data=jci_jif[which(jci_jif$Indicator=='5-Year Journal Impact Factor'),], aes(fill=Correlation, size=Docs), shape=21, color='black', stroke=0.5, alpha=0.6)+
  scale_fill_gradient(low='white', high='blue', na.value=NA, breaks=c(0, 0.5, 1), limits=c(0,1))+
  scale_size_continuous(range=c(1,5), breaks=c(0, 25000,50000, 100000, 125000, 150000, 200000))+
  theme_light()+
  theme(panel.grid=element_blank(),
        text=element_text(family='Arial', size=12.5, color='black'),
        axis.text=element_text(color='black', size=11),
        axis.ticks=element_line(color='black'),
        legend.position='none',
        legend.box='vertical',
        panel.border=element_rect(colour='black'),
        strip.background=element_rect(colour='black', fill='black'),
        strip.text = element_text(size=14))+
  guides(fill=guide_colorbar(label.position='bottom',
                             title.position='left', title.vjust=0.85)) +
  labs(x='')+
  facet_wrap(~Indicator)
```

## 2.2. AIS, Eigenfactor, Immediacy Index and Total Citations
The correlations between the JCI and the rest of the indicators are calculated, but taking into account only those categories with more than 25 journals. This is because a few small categories greatly alter the results.
```{r other_correlations_df}
ot_indicators <- df_cor[which((!(df_cor$Indicator %in% c('2020 JIF', '5 Year JIF'))) & df_cor$Journals>25),]

ot_indicators[which(ot_indicators$DB=='SSCI'),'DB'] <- 'Social\nSciences'
ot_indicators[which(ot_indicators$DB=='SCIE'),'DB'] <- 'Science'
ot_indicators[which(ot_indicators$DB=='AHCI'),'DB'] <- 'Arts &\nHumanities'
ot_indicators[which(ot_indicators$DB=='ESCI'),'DB'] <- 'Emerging\nSources'
```

There are 118 correlations with a p-value less than 0.01 (43 in Total Citations, 33 in Immediacy Index and Eigenfactor, and 9 in Article Influence Score).
```{r cor_test}
table(ot_indicators[which(ot_indicators$p >= 0.01),'Indicator'])
```

Boxplots are generated.
```{r other_correlations, fig.height=7, fig.width=9, message=FALSE}
ggplot(data=ot_indicators, aes(y=Correlation, x=DB, group=DB))+
  geom_boxplot(outlier.shape=NA, width=0.8) +
  geom_jitter(aes(fill=Correlation, size=Docs), shape=21, color='black', stroke=0.5, alpha=0.6)+
  scale_fill_gradient(low='white', high='#eb2701', na.value=NA, breaks=c(0, 0.5, 1), limits=c(0,1))+
  scale_size_continuous(range=c(0.1,3.5), breaks=c(0, 25000,50000, 100000, 125000, 150000, 200000))+
  theme_light()+
  theme(panel.grid=element_blank(),
        text=element_text(family='Arial', size=12.5, color='black'),
        axis.text=element_text(color='black', size=10.5),
        axis.ticks=element_line(color='black'),
        legend.position='bottom',
        panel.border=element_rect(colour='black'),
        strip.background=element_rect(colour='black', fill='black'),
        strip.text=element_text(size=14))+
  guides(fill = guide_colorbar(label.position='bottom',
                               title.position='left', title.vjust=0.85)) +
  labs(x='')+
  facet_wrap(~Indicator)
```

# 3. Regression analysis
Firstly, we rank JIF and JCI values by categories.
```{r to_ranks}
df_rank <- df[which(!(is.na(df$`2020 JIF`) | is.na(df$`2020 JCI`))),]

length(unique(df_rank$Category))

df_rank$quantile_jif <- NA
df_rank$quantile_jci <- NA

for(ca in unique(df$Category)){
  df_rank[which(df_rank$Category==ca),'quantile_jif'] <- trunc(rank(-df_rank[which(df_rank$Category==ca),'2020 JIF'], ties.method = 'random'))
  df_rank[which(df_rank$Category==ca),'quantile_jci'] <- trunc(rank(-df_rank[which(df_rank$Category==ca),'2020 JCI'], ties.method = 'random'))
}
```

A linear regression model is generated with JIF and JCI ranks.
```{r general_lm}
lm_g <- lm(quantile_jif~quantile_jci, df_rank[which(df_rank$DB %in% c('SSCI', 'SCIE')),])
summary(lm_g)
```
The models obtained for SSCI and SCIE journals have a good R^2^.
```{r linear_models}
lm_ssci <- lm(quantile_jif~quantile_jci, df_rank[which(df_rank$DB == 'SSCI'),])
summary(lm_ssci)$r.squared

lm_scie <- lm(quantile_jif~quantile_jci, df_rank[which(df_rank$DB == 'SCIE'),])
summary(lm_scie)$r.squared
```

```{r linear_models_plot, dpi=200, fig.retina=2, fig.height=5, fig.width=9, warning=FALSE, message=FALSE}
df_r2 <- data.frame(quantile_jci=c(Inf,Inf),
                    quantile_jif=c(10,10),
                    text=c(paste0('y==', format(coef(lm_ssci)[1], digits = 2), '+', format(coef(lm_ssci)[2], digits = 2), '~x~', '~~italic(r)^2==', round(summary(lm_ssci)$r.squared,3),''),
                           paste0('y==', format(coef(lm_scie)[1], digits = 2), '+', format(coef(lm_scie)[2], digits = 2), '~x~', '~~italic(r)^2==', round(summary(lm_scie)$r.squared,3),'')),
                    DB=c('SSCI','SCIE'))

ggplot(data=df_rank[which(df_rank$DB %in% c('SSCI', 'SCIE')),], aes(x=quantile_jci, y=quantile_jif)) +
  geom_point(color='#1f77b4', size=1, alpha=0.75, stroke=0)+
  geom_smooth(formula = 'y ~ x', method='lm', color='black')+
  geom_text(data=df_r2, aes(x=quantile_jci, y=quantile_jif, label=text, group=DB),
            parse = TRUE, hjust=1.05) +
  theme_light() +
  labs(x='JCI rank', y='JIF rank') +
  facet_wrap(.~DB, scales = 'free') +
  theme(text=element_text(family='Arial', size=12.5, color='black'),
        axis.text=element_text(color='black', size=11),
        axis.ticks=element_line(color='black'),
        panel.border=element_rect(color='black'),
        strip.background=element_rect(color='black', fill='black'),
        strip.text=element_text(size=12, color='white'),
        legend.position='none'
  ) 
```

The average R^2^ value of the linear regression models is calculated.
```{r average_lm}
df_r <- data.frame(category=unique(df_rank$Category[which(df_rank$DB %in% c('SSCI', 'SCIE'))]),
                   r2=NA,
                   DB=NA,
                   stringsAsFactors = FALSE)

for(ca in unique(df_r$category)){
  lm_aux <- NA
  lm_aux <-lm(quantile_jci~quantile_jif,df_rank[which(df_rank$Category==ca & !is.na(df_rank$`2020 JIF`) & !is.na(df_rank$`2020 JCI`)),])
  df_r[which(df_r$category==ca),'r2'] <- summary(lm_aux)$r.squared
  df_r[which(df_r$category==ca),'DB'] <- unique(df_rank[which(df_rank$Category==ca),'DB'])
}

mean(df_r$r2, na.rm = TRUE)
mean(df_r$r2[which(df_r$DB=='SSCI')], na.rm = TRUE)
mean(df_r$r2[which(df_r$DB=='SCIE')], na.rm = TRUE)
```

## 3.1 Information Science & Library Science model
As specific study case, the linear regression model for Information Science & Library Science is generated.
```{r}
df_lib <- df[which(df$Cat == 'INFORMATION SCIENCE & LIBRARY SCIENCE'),]
df_lib_train <- df_lib[which(!(is.na(df_lib$`2020 JIF`) | is.na(df_lib$`2020 JCI`))),]

df_lib_train$quantile_jif <- trunc(rank(-df_lib_train[,'2020 JIF'], ties.method = 'first'))
df_lib_train$quantile_jci <- trunc(rank(-df_lib_train[,'2020 JCI'], ties.method = 'first'))

lib_lm <- lm(quantile_jif~quantile_jci, df_lib_train)
```

JIF ranking is predicted and compared to the original JIF ranking.
```{r prediction, dpi=200, fig.retina=2, fig.height=5, fig.width=9, warning=FALSE, message=FALSE}
df_lib_predict <- df_lib[which(!(is.na(df_lib$`2020 JCI`))),]
df_lib_predict$quantile_jci <- trunc(rank(-df_lib_predict[,'2020 JCI'], ties.method = 'first'))
df_lib_predict$quantile_jif <- trunc(rank(-df_lib_predict[,'2020 JIF'], ties.method = 'first'))
df_lib_predict$quantile_jif[which(is.na(df_lib_predict$`2020 JIF`))] <- NA

df_lib_predict$quantile_jci_p <- predict(lib_lm, newdata = df_lib_predict)
df_lib_predict$quantile_jci_p <- rank(as.numeric(df_lib_predict$quantile_jci_p), ties.method = 'random')

ggplot() +
  geom_point(data=df_lib_predict, aes(x='JIF', y=quantile_jif, color=DB), alpha=0.7) +
  geom_point(data=df_lib_predict, aes(x='Predicted JIF', y=quantile_jci_p, color=DB), alpha=0.7) +
  geom_segment(data =df_lib_predict, aes(x = 'Predicted JIF', xend = 'JIF', y = quantile_jci_p, yend = quantile_jif),
               inherit.aes = FALSE,
               color='grey20',
               size=0.25) +
  geom_text(aes(y=c(1,dim(df_lib_predict)[1]/4, 2*dim(df_lib_predict)[1]/4, 3*dim(df_lib_predict)[1]/4),
                x=Inf,
                label=c('Q1','Q2','Q3','Q4')),
            hjust=1.2, vjust=1)+
  scale_color_manual(values=c('#fc4e08', '#28b7c1'), labels = c('ESCI journals', 'SSCI journals'))+
  geom_hline(yintercept = c(dim(df_lib_predict)[1]/4, 2*dim(df_lib_predict)[1]/4, 3*dim(df_lib_predict)[1]/4), size=0.25)+
  coord_flip()+
  scale_x_discrete(limits=rev)+
  scale_y_continuous(trans = "reverse", breaks = c(1,50,100,150)) +
  labs(title = '', y='Ranking position', x='')+
  theme_classic()+
  theme(text=element_text(family='Arial', size=12.5, color='black'),
        axis.text=element_text(color='black', size=11),
        axis.ticks=element_line(color='black'),
        legend.position = 'bottom',
        legend.title = element_blank())
```